A pipeline to analyse time-course gene expression data

### 2020-06-17 scripts/manuscript.Rmd
Continuing some small fixes here and there

# Now use the results of the previous call to aheatmap to reorder the genes.
aheatmap(
    moanin:::rescale_values(data_to_plot)[res$rowInd,],
    Colv=NA,
    Rowv=NA,
    annCol=submeta[,
                c("Group", "Timepoint")],
    annLegend=TRUE,
    annColors=ann_colors,
    main=paste("Cluster", cluster_to_plot, "(rescaled)"),
    treeheight=0)

@@

**EAP:** We shouldn't use any hidden functions (i.e. that use `:::` as in `moanin:::rescale_values`). We should make it a exported function if use it in workflow.

@@

**EAP** This didn't make sense to me: "run the clustering for all possible clusters of interest." Do you mean across different values of $k$? That's what I changed it to.

@@

**EAP**: Why don't you use the same rule/code as above? i.e. with the fisher p-values etc.? A little strange to have two different filtering rules for the clustering. Or is this the same, and you're restating the rule? (Unclear why you'd have to redo everything here...didn't go back and compare code)

```{r bootstrapped_clustering, eval=FALSE}
pvalues = de_analysis[, pval_col_to_keep]

splines_model = moanin::create_moanin_model(meta, degrees_of_freedom=6)
fishers_pval = moanin:::pvalues_fisher_method(pvalues)
fishers_qval = stats::p.adjust(fishers_pval)

# Keep only the genes found differentially expressed after combining the
# p-values using Fisher's method and with a log fold change greater than 2
genes_to_keep = row.names(
    lfc_max[(rowSums(lfc_max > 2) > 0) & (fishers_qval < 0.05), ])
y = as.matrix(data[genes_to_keep, ])

# You may want to set the random seed of the experiment so that the results
# don't vary if you rerun the experiment.
# set.seed(random_seed)
n_genes = dim(y)[1] * sample_proportion
indices = sample(1:dim(y)[1], n_genes, replace=TRUE)

@@

Now we will use the function `consensus_matrix` in the `moanin` package to
calculate proportion of times each pair of samples was clustered together
across the 20 resampling runs, and plot the heatmap of those consensus
matrices for $k=5$ and $k=20$.

**EAP** It's somewhat unfortunately that B=20 and the maximum k=20. Makes it easy to get them confused. Could we switch one of these to be 15 or 25 so as to keep them straight for the reader?

@@

```{r echo=FALSE, results="hide"}
n_clusters = 2:20
all_labels = list()
for(i in n_clusters){
    filename = paste0("results/stability_", i, ".tsv")
    stability = read.table(filename, sep="\t")
    all_labels[[paste0("B", i)]] = stability
}
```

@@

```{r plotModelExplorer, fig.width=6, fig.height=6,fig.cap="Plot of the cumulative density of similarity of clusterings of data for different k"}
plot_model_explorer(all_labels)
```
**EAP** I'm confused by this plot. Isn't the numbering here for different values of $k$? Why do the labels say "B1" and "B2"? Isn't that the notation above for the bootstrap replications for each values of $k$? It seems like the names of `all_labels` should be something like `k=1`,..., while the names of `all_labels[[1]]` should be `B1,B2,...`. But currently they have the same values.

@@

```{r  plotModelExplorerZoom, fig.width=6, fig.height=6,fig.cap="Plot of the cumulative density of similarity of clusterings of data for different k, restricted to k=6-10"}
clusters = c("B6", "B7", "B8", "B9", "B10")
selected_labels = all_labels[clusters]
plot_model_explorer(selected_labels)
```

>>>>

# Now use the results of the previous call to aheatmap to reorder the genes.
aheatmap(
    moanin::rescale_values(data_to_plot)[res$rowInd,],
    Colv=NA,
    Rowv=NA,
    annCol=submeta[,
                c("Group", "Timepoint")],
    annLegend=TRUE,
    annColors=ann_colors,
    main=paste("Cluster", cluster_to_plot, "(rescaled)"),
    treeheight=0)

@@

@@

@@

```{r bootstrapped_clustering, eval=FALSE}
# You may want to set the random seed of the experiment so that the results
# don't vary if you rerun the experiment.
# set.seed(random_seed)
n_genes = dim(y)[1] * sample_proportion
indices = sample(1:dim(y)[1], n_genes, replace=TRUE)

@@

Now we will use the function `consensus_matrix` in the `moanin` package to
calculate proportion of times each pair of samples was clustered together
across the 25 resampling runs, and plot the heatmap of those consensus
matrices for $k=5$ and $k=20$.

@@

```{r echo=FALSE, results="hide"}
n_clusters = 2:20
all_labels = list()
for(i in n_clusters){
    filename = paste0("results/stability_", i, ".tsv")
    stability = read.table(filename, sep="\t")
    all_labels[[paste0("k=", i)]] = stability
}
```

@@

```{r plotModelExplorer, fig.width=6, fig.height=6,fig.cap="Plot of the cumulative density of similarity of clusterings of data for different k"}
plot_model_explorer(all_labels)
```

@@

```{r  plotModelExplorerZoom, fig.width=6, fig.height=6,fig.cap="Plot of the cumulative density of similarity of clusterings of data for different k, restricted to k=6-10"}
clusters = c("k=6", "k=7", "k=8", "k=9", "k=10")
selected_labels = all_labels[clusters]
plot_model_explorer(selected_labels)
```
